% vim: set spell:

\section{Recommendations for Amazon}

\begin{enumerate}[a.]

\item

\begin{enumerate}[1.]

\item Amazon.com's product recommendations can be classified as a form of query
term suggestions \cite[\textsection\ 1.5.4]{hearst}. The information need here
is less explicit. When a user follows recommendations, it is only sometimes to
fulfill the original information need. Recommendations offer similar material
that may also be of interest. They fulfill an information need which the user
was not necessarily consciously aware of.

\item Amazon.com's product reviews do not immediately fit any of the
recommendations for search user interface design presented in the literature.
One way that Amazon.com could improve upon these is to allow users to add tags
to products, so that richer networks of categories of products can emerge. This
would also provide information for the search engine, which would otherwise
require more complicated processing of user reviews.

\end{enumerate}

\item

The suggestions below are enumerated in correspondence with the problems
identified in exercise 1.

\begin{enumerate}[1.]

\item Let a user pick a default shipping address. Consider this address before
presenting an item as either a search result, or presenting a buy or a wish
list button. If the item cannot be delivered to the default shipping address,
do not show it in search results, and instead of a buy or an add to wish list
button, inform the user of the fact that this item cannot be delivered. It
might however be of benefit to let users pick a geographical region of
delivery, so e.g. they can compile a wish list for their distant relatives.

This way, a user is never let astray believing that they can obtain an item,
when they indeed cannot.

\item Consolidate all the web sites into one. In search results, present only
those items that can be delivered to the geographical regions that a user has
picked.It may be possible to order the same item from different locations (e.g.
United States, United Kingdom, Germany). Let the user pick the one that leads
to cheapest delivery price and lowest delivery time.

This way, a user is never let astray believing that they can obtain an item,
when they indeed cannot.

This option is extremely costly as it demands a rework of the entire Amazon.com
brand. A simpler option would be to give AmazonGlobal a more clear presence on
the web. This somewhat reduces the risk of letting the users astray.

\item Show the total amount to be paid when or before asking for credit card
information.  Although the home currency of the buyer may not revealed yet, the
buyer is fully aware of the currency used to price the item they are now
attempting to buy.

\item Make the distinction more clear, perhaps with the use of background
colours, whether a book is physical or an Amazon Kindle book. Do this both in
search results and on the product pages. The distinction is important enough to
deserve this treatment, as a reader without an Amazon Kindle reader cannot read
an Amazon Kindle book, and would feel mislead having purchased such a book.

\item Put the shipping rates and times together, or at least have them no more
than back-click and hyperlink away from one another. This way the users will
find delivery information more quickly, and proceed with (or quit) shopping.

\item Do not let items that cannot be delivered to the user's chosen
geographical region be added to their wish list. This way, the users won't
build up wish lists of items they (and most likely their family or friends)
cannot buy. Consider disallowing compiling wish lists composed of items that
cannot all be delivered to a particular geographical region.

\item Don't save items removed from the shopping cart for later purchases, when
these items cannot be delivered to the user's chosen geographical region.

\item Let the pictorial representation adhere to the order of elements in the
search results when sorted by rating, or vice versa. This will avoid users
getting confused when seeming lower rated items precede higher rated items when
sorted by nondecreasing rating.

\end{enumerate}

\item The below are inspired by \cite[\textsection\ 1]{hearst}.

\begin{itemize}

\item Make it easier for users to recover from errors and avoid them
altogether. For instance, make sure that it's not possible to place an item
into the shopping cart which cannot be shipped to their chosen geographical
region. Removing an item from the shopping cart is a laborious process,
consider a step-counter alongside the text box.

\item Inform the users more clearly of what is going on. The user tests showed
problems with seeing the notifications that the items could not be delivered to
the particular address, and switching to Kindle versions of books rather than
physical. See also problem 7.

\item Suggest that the user makes an explicit choice when searching for book
titles, whether they would like to see results for Amazon Kindle books or
physical books.

\end{itemize}

\end{enumerate}

